---
title: "Data visualization workshop: basics in ggplot2"
author: "Katie Florko"
date: "June 16, 2017"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
---

***
# Introduction

#### Getting started

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#load packages
library(ggplot2)
```


***
#### Getting started with data
We will use a made up dataset on polar bears
```{r}
data <- read_csv("bears.csv")
head(data, 5)
```


# Introduction to formatting
```{r}
#a basic plot
ggplot(data, aes(Year, Seal)) +geom_point()

#different themes
ggplot(data, aes(as.factor(Year), Seal)) +geom_point()+theme_bw()
ggplot(data, aes(as.factor(Year), Seal)) +geom_point()+theme_classic()
ggplot(data, aes(as.factor(Year), Seal)) +geom_point()+theme_grey()

#change axis labels
ggplot(data, aes(as.factor(Year), Seal)) +geom_point()+theme_grey()+labs(x = "Year",y ="Mean proportion of seal in polar diet (%)") 

#split up text
ggplot(data, aes(as.factor(Year), Seal)) +geom_point()+theme_grey()+labs(x = "Year",y ="Mean proportion\nof seal in polar diet (%)") 

#add title
ggplot(data, aes(as.factor(Year), Seal)) +geom_point()+theme_grey()+labs(x = "Year",y ="Mean proportion\nof seal in polar diet (%)") + ggtitle("A nice title")

#center title
ggplot(data, aes(as.factor(Year), Seal)) +geom_point()+theme_grey()+labs(x = "Year",y ="Mean proportion\nof seal in polar diet (%)") + ggtitle("A nice title")+theme(plot.title = element_text(hjust = 0.5))

#color by sex (a factor)
ggplot(data, aes(as.factor(Year), Seal)) +geom_point(aes(color=Sex))+theme_grey()+labs(x = "Year",y ="Mean proportion\nof seal in polar diet (%)") + ggtitle("A nice title")+theme(plot.title = element_text(hjust = 0.5))


#color by gear (a continuous variable)
ggplot(data, aes(as.factor(Year), Seal)) +geom_point(aes(color=Bodycond))+theme_grey()+labs(x = "Year",y ="Mean proportion\nof seal in polar diet (%)")+ ggtitle("A nice title")+theme(plot.title = element_text(hjust = 0.5))

#modify point size
ggplot(mtcars, aes(wt, mpg))+ geom_point(aes(size = disp, colour=factor(gear)))
```
# Other types of plots 

## Box plots
We need factoral data to make a boxplot. In the code, we will refer to gears and cyl as factors (i.e., as.factor(mtcars$gear)), but if you are doing this with your own data, the variable you will want to look via a boxplot should already be a factor
```{r}
as.factor(mtcars$gear)
ggplot(mtcars, aes(x = as.factor(mtcars$gear), y = wt, group =as.factor(mtcars$cyl))) +geom_boxplot()

data <- read_csv("bears.csv")
```

## Line Graphs
```{r}

```

## Bar Graphs
```{r}

```

## Maps
```{r}

```

# Facet grids and wraps
```{r}

```

