---
title: "Data visualization workshop: making maps using ggplot2 and ggmap"
author: "Katie Florko"
date: "June 16, 2017"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
---

***
# Introduction

#### Getting started

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#load packages
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggmap)
```


***
#### Getting started with data
We will use a made up dataset on polar bears
```{r}
setwd("~/Documents/R/YorkU.DataViz.2017-06-19/data")
data <- read.csv("bears.csv")
head(data, 5)
```

#First, use ggplot2
```{r}
world<-map_data("world","Canada")
w <- fortify(world)

map <-ggplot(data, aes(Long,Lat)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat))+coord_map()+ ylim(67.5, 76.5) + xlim(-142,-100)

map + geom_point(data=data, aes(Long, Lat))

map + geom_point(data=data, aes(Long, Lat, shape=Sex, color=Sex),size=3)+  scale_color_manual(name="Sex", values=c("red","black")) + scale_shape_manual(name="Sex", values=c(15:16))
```

#Change color/shape of variables: Discrete variable
```{r}
#Basic
ggplot(data, aes(Long,Lat)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat))+coord_map()+ ylim(67.5, 76.5) + xlim(-142,-100)+ geom_point(data=data, aes(Long, Lat))

#Color by ageclass
ggplot(data, aes(Long,Lat, color=Ageclass)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat))+coord_map()+ ylim(67.5, 76.5) + xlim(-142,-100)+ geom_point(data=data, aes(Long, Lat))

#Manually select colors
ggplot(data, aes(Long,Lat, color=Ageclass)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat))+coord_map()+ ylim(67.5, 76.5) + xlim(-142,-100)+ geom_point(data=data, aes(Long, Lat))+  scale_color_manual(name="Ageclass", values=c("red","black","green"))

#Shape and color by ageclass 
ggplot(data, aes(Long,Lat, color=Ageclass)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat))+coord_map()+ ylim(67.5, 76.5) + xlim(-142,-100)+ geom_point(data=data, aes(Long, Lat, shape=Ageclass))+  scale_color_manual(name="Ageclass", values=c("red","black","green"))+ scale_shape_manual(name="Ageclass", values=c(15:17))
```

#Change color/size of variables: Continuous variable
```{r}
#Basic
ggplot(data, aes(Long,Lat)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat))+coord_map()+ ylim(67.5, 76.5) + xlim(-142,-100)+ geom_point(data=data, aes(Long, Lat))

#Color by Bodycond
ggplot(data, aes(Long,Lat, color=Bodycond)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat))+coord_map()+ ylim(67.5, 76.5) + xlim(-142,-100)+ geom_point(data=data, aes(Long, Lat))


#Color by Bodycond: manually set limits and colors
ggplot(data, aes(Long,Lat, color=Bodycond)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat))+coord_map()+ ylim(67.5, 76.5) + xlim(-142,-100)+ geom_point(data=data, aes(Long, Lat))+scale_color_gradient(limits=c(0,100), low = "green", high = "black", name = "A nice\nnew title")

#Color by Bodycond: break up gradient legend
ggplot(data, aes(Long,Lat, color=Bodycond)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat))+coord_map()+ ylim(67.5, 76.5) + xlim(-142,-100)+ geom_point(data=data, aes(Long, Lat))+scale_color_continuous(limits=c(40,100), low = "green", high = "black", name = "Body\ncondition",breaks = c(45, 95), labels = c("Skinny", "Fat"))

#Size by Bodycond: break up gradient legend
ggplot(data, aes(Long,Lat, color=Bodycond)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat))+coord_map()+ ylim(67.5, 76.5) + xlim(-142,-100)+ geom_point(data=data, aes(Long, Lat, size=Walrus))+scale_color_continuous(limits=c(40,100), low = "green", high = "black", name = "Body\ncondition",breaks = c(45, 95), labels = c("Skinny", "Fat"))
```

#Add text
```{r}
#Basic
ggplot(data, aes(Long,Lat)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat))+coord_map()+ ylim(67.5, 76.5) + xlim(-142,-100)+ geom_point(data=data, aes(Long, Lat))

#Add title
ggplot(data, aes(Long,Lat)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat))+coord_map()+ ylim(67.5, 76.5) + xlim(-142,-100)+ geom_point(data=data, aes(Long, Lat))+labs(title="New title")

#Add place names
ggplot(data, aes(Long,Lat)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat))+coord_map()+ ylim(67.5, 76.5) + xlim(-142,-100)+ geom_point(data=data, aes(Long, Lat))+labs(title="New title")+geom_text(aes(x=-127, y=68.5, label="N.W.T."), size=3,color = "black")+geom_text(aes(x=-135, y=75, label="Beaufort\nSea"), size=4, color = "darkblue", fontface = "italic")+geom_text(aes(x=-110, y=71, label="Victoria\nIs."), size=3,color = "black")+geom_text(aes(x=-121, y=73, label="Banks\nIs."), size=3, color = "black")+geom_text(aes(x=-111.5, y=75.75, label="Viscount\nMelville Is."), size=3, color = "black") 
```

#Ellipses
```{r}
#Add ellipse to all points
map <-ggplot(data, aes(Long,Lat)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat))+coord_map()+ ylim(67.5, 76.5) + xlim(-142,-100)+geom_point(data=data, aes(Long, Lat, shape=Sex, color=Sex))+  scale_color_manual(name="Sex", values=c("red","blue")) + scale_shape_manual(name="Sex", values=c(15:16))

l<- map+geom_point(data=data, aes(x=Long, y=Lat, shape=Sex,color = Sex))+
stat_ellipse(geom="polygon", fill=NA,color="black",level = 0.50, type="t", size=2)+
stat_ellipse(geom="polygon", fill=NA, color="black",level = 0.75, type="t", linetype=2)
l

#Add ellipse, based on factor variable
map<-ggplot(data, aes(Long, Lat, color = Sex)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat, group=group)) + xlab("Longitude") +ylab("Latitude") + ylim(67.5, 76.5) + xlim(-142,-100) +coord_map() +guides(fill = guide_legend(reverse = TRUE))

l <- map  + geom_point(data=data, aes(x=Long, y=Lat, shape=Sex,color = Sex)) + theme_bw()  + coord_map() +guides(fill = guide_legend(reverse = TRUE))+
stat_ellipse(geom="polygon", fill=NA, level = 0.50, type="t", size=2)+
stat_ellipse(geom="polygon", fill=NA, level = 0.75, type="t", linetype=2)+ ggtitle("All1, all bears, all FA")+  scale_color_manual(name="Sex", values=c("blue","red")) + scale_shape_manual(name="Sex", values=c(15:16))
l
```

#Add heat map

```{r}
map <- ggplot(data, aes(Long,Lat)) + geom_polygon(data=w, colour="grey50", fill="grey50", aes(x=long, y=lat))+coord_map()+ ylim(67.5, 76.5) + xlim(-142,-100)+ geom_point(data=data, aes(Long, Lat))

heat <- map +stat_density2d(aes(fill=..level..,alpha=..level..),geom='polygon',colour=NA, na.rm = TRUE, bins=20,h=10) + scale_fill_continuous(low="green",high="red")  +
labs(fill = "Sampling\nIntensity")  + guides(alpha="none") +geom_point(na.rm=TRUE) + theme(legend.position = c(0.14, 0.65),panel.grid.major = element_line(colour = "white"))
heat

#change number of bands (bandwidth = bins) and number of grid points in each direction (n)
heat <- map +stat_density2d(aes(fill=..level..,alpha=..level..),geom='polygon',colour=NA, na.rm = TRUE, bins=5,h=9) + scale_fill_continuous(low="green",high="red")  +
labs(fill = "Sampling\nIntensity")  + guides(alpha="none") +geom_point(na.rm=TRUE)
heat

#while we're here, change location of legend 
heat <- map +stat_density2d(aes(fill=..level..,alpha=..level..),geom='polygon',colour=NA, na.rm = TRUE, bins=5,h=9) + scale_fill_continuous(low="green",high="red")  +
labs(fill = "Sampling\nIntensity")  + guides(alpha="none") +geom_point(na.rm=TRUE) + theme(legend.position = c(0.1, 0.7),panel.grid.major = element_line(colour = "white"))
heat

#and change background to transparent
heat <- map +stat_density2d(aes(fill=..level..,alpha=..level..),geom='polygon',colour=NA, na.rm = TRUE, bins=5,h=9) + scale_fill_continuous(low="green",high="red")  +
labs(fill = "Sampling\nIntensity")  + guides(alpha="none") +geom_point(na.rm=TRUE) + theme(legend.position = c(0.1, 0.7),panel.grid.major = element_line(colour = "white"), legend.background = element_rect(fill="NA"))
heat
```

#GGMAP
The only difference here is that you can "get_map" and formatting is set up to automatically make it a map (without having to call it a geom_polygon)
```{r}
geocode("york university")
geocode("cn tower", output="all")

qmap(location="northamerica", zoom=5)
qmap(location="york university")
qmap(location="york university", zoom = 14, extent="normal")

qmap(location="northamerica", zoom=3)
```

##plot crimes in houston
```{r}
# first, subset to have only violent crimes
violent_crimes <- subset(crime,offense != "auto theft" & offense != "theft" & offense != "burglary")

# subset area to downtown
violent_crimes <- subset(violent_crimes,
  -95.39681 <= lon & lon <= -95.34188 &
   29.73631 <= lat & lat <=  29.78400)

#make basemap
HoustonMap <- qmap("houston", zoom = 14,color = "bw")

#add layer with crime location data
HoustonMap +geom_point(aes(x = lon, y = lat),data = violent_crimes)

#Color by type of crime
HoustonMap +geom_point(aes(x = lon, y = lat,colour = offense),data = violent_crimes)
```

#another heat map example
```{r}
#make basemap
houston <- get_map(location = "houston", zoom = 14, color = "bw")
HoustonMap <- ggmap(houston, base_layer = ggplot(aes(x = lon, y = lat), data = violent_crimes))

#add crimes (stat_density2d) and by offense (facet_wrap)
HoustonMap + stat_density2d(aes(x = lon, y = lat, fill = ..level.., alpha= ..level..),
bins = 5, geom = "polygon",
data = violent_crimes) +scale_fill_gradient(low = "green",high= "red") +facet_wrap(~ month)

```


